---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

##section 1. The PDB
# determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein? 
```{r}
stats <- read.csv("Data Export Summary.csv")
head(stats)
```

```{r}
sum(stats$Total)
```
```{r}
percent.by.method <- stats$Total/sum(stats$Total) * 100
names(percent.by.method) <- rownames(stats)
percent.by.method
```

```{r}
sum(stats$Proteins)/ sum(stats$Total) * 100
```

Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

```{r}
library(bio3d)
```

now we can use the fcs from the bio3d package

```{r}
pdb <- read.pdb("1hsg")
print(pdb)
```

```{r}
attributes(pdb)
```
```{r}
head(pdb$atom)
```

```{r}
# Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```

```{r}
# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]
```

```{r}
# Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab=“B-factor”)
```

```{r}
atom.select(pdb, )
pro
```


```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```


```{r}
# Print details of the first few selected atoms
head( pdb$atom[ca.inds$atom, ] )
```

```{r}
# And selected xyz coordinates
head( pdb$xyz[, ca.inds$xyz] )
```



```{r}
# Select chain A
a.inds <- atom.select(pdb, chain="A")
```

```{r}
# Select C-alphas of chain A
ca.inds <- atom.select(pdb, "calpha", chain="A")
```

```{r}
# We can combine multiple selection criteria to return their
#intersection
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A",
resno=10:20)
```

#we want to select out the protein and deug only parts o
```{r}
prot.inds <- atom.select(pdb, "protein")
prot.inds
```

```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file="protein.pdb")
```

#let's do the same thing for our drug

```{r}
lig.inds <- atom.select(pdb, "ligand")
lig.pdb <- trim.pdb(pdb, lig.inds)
write.pdb(lig.pdb, file="ligand.pdb")
```

```{r}
install.packages("devtools")
```


```{r}
library(bio3d)
View(pdb, "overview")
```

