---
title: "class 15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#import coldata and metadata
```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
dim(counts)
head(counts)
```


```{r}
head(metadata)
```


```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```

```{r}
#meancounts <- data.frame(control.mean, treated.mean)
```

First task is to find which columns in **counts** correspond to the control conditions (i.e. no drug!) the we need to calculate the mena count value for each gene. 

```{r}
control.inds <- metadata$dex == "control" 
control.ids <- metadata$id[control.inds]

head( counts[,control.ids] )

```
calculated the average COUNTS value for all genes
```{r}
rowSums(counts[,control.ids]) / ncol(counts[,control.ids])
```


Now do the same thing but for the experimental groups//drug. Determine which columns corresspond to drug treated groups and then calculate the average value of each gene. (** i.e. drug treated groups**)

```{r}
treated.inds <- metadata$dex == "treated"
treated.ids <- metadata$id[treated.inds]

head( counts[, treated.ids] )
```
calculate the avg. of COUNT value for drug treated genes
```{r}
 treated.mean <- rowSums(counts[,treated.ids]) /ncol(counts[,treated.ids])
```

combine the mean count for all genes from the control and treated

```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

```{r}
colSums(meancounts)
```

Now we can plot the drug treated vs. non-drug control count values
```{r}
plot(meancounts$control.mean,meancounts$treated.mean)
```
take the log

```{r}
plot(log2(meancounts$control.mean),log2(meancounts$treated.mean))
```

We can find candidate differentially expressed genes by looking for genes with a large change between control and dex-treated samples. We usually look at the log2 of the fold change, because this has better mathematical properties.

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or the View() function for example.

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
the which** functoin shows us which coulumn, variable, etc. is true/false to the statement.
```{r}
ty <- c(1,2,4,56,7,0,8,0)
ty ==0
which(ty==0)
```


```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
ty <- meancounts[1:6,]
ty[,1:2]
unique( which(ty[,1:2] ==0, arr.ind=TRUE)[,1])
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
up.ind
```


```{r}
sum(up.ind)
```

```{r}
sum(down.ind)
```

there are 'sum(up.ind)' up regulated genes and 'sum(down.ind)' down regulated genes. 


all of these gene are upregulated...but nOW what do we do? return to biology speak and determine what each gene does and how it connects to asthma. 
```{r}
head(mycounts[up.ind,])
```


use the merge() function to add annotation data from ensemlbe


```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```


```{r}
mycounts.annotated <- merge(x=mycounts, y=anno, by.x="row.names", by.y = "ensgene")
head(mycounts.annotated)
```



#DEseq2 analysis

```{r}
library(DESeq2)

```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```

```{r}
summary(res)
```

```{r}
plot(res$log2FoldChange, -log(res$padj))
```






