---
title: "class06"
author: "Felipe Rangel"
date: "1/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section 1: Reading files again :-)


Here we are trying to use the **read.table()** function and friends to read some example flat files. 
First let's try adding a simple plot

```{r}
plot(1:10)
```

Back to file reading...
```{r}
read.table("https://bioboot.github.io/bimm143_W19/class-material/test1.txt", header= TRUE, sep=",")
```

for this commmon CSV format we can use **read.csv**

```{r}
file1 <- "https://bioboot.github.io/bimm143_W19/class-material/test1.txt"
data1 <- read.csv(file1)
data1
```
for this common file, i am trying to figure out how to read this file using the different functions we've learned so far.

```{r}
file2 <-"https://bioboot.github.io/bimm143_W19/class-material/test2.txt"
data2 <- read.csv2(file2, header = TRUE, sep = "$")
data2
```

for this next common file, I need to repeat the same steps as in for the command above...

```{r}
file3 <- "https://bioboot.github.io/bimm143_W19/class-material/test3.txt"
data3 <- read.table(file3)
data3
```

## Section 2. R funcitons

My first function

```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

let's try using the function
```{r}
add(1)

```

```{r}
add(1,100)
```

```{r}
add( c(1, 2, 3) )
```
```{r}
add( c(1, 2, 3), 4 )
```

```{r}
#add(1, 2, 2)
```
```{r}
#add(x=1, y="b")
```


my 2nd function

```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1]) 
}
```

Test on a small sample, where we know what the answ should be
```{r}
rescale(1:10)
```

```{r}
# How would you get your function to work here…
rescale( c(1,2,NA,3,10) )
```

```{r}
# What should your function do here?
# rescale( c(1,10,"string") )
```

```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm=TRUE)
 (x - rng[1]) / (rng[2] - rng[1]) 
}
x <- c(1,2,NA,3,10)
rng <- range(x)
rng
```
```{r}
rescale2( c(1,2,NA,3,10))
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 if(na.rm) {
 rng <-range(x, na.rm=TRUE)
 } else {
 rng <-range(x)
 }
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("is it me you are looking for?")
 
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```


```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 if(na.rm) {
 rng <-range(x, na.rm=TRUE)
 } else {
 rng <-range(x)
 }
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 return(answer)
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
}

```


